---
lang: en
title: "PRIMED iCite report "
author: "PRIMED CC"
date: "`r lubridate::today()`"
format:
  html:
    toc: true
    self_contained: yes
execute:
  echo: false
output:
  rmdformats::downcute:
    fig_width: 10
    fig_height: 4
    thumbnails: false
    css: "style.css"
params:
    records_file: "tmp/icite_records.json"
    search_id_file: "tmp/icite_search_id.txt"
---


```{r library, include=FALSE}
library(tidyverse)
library(knitr)
library(jsonlite)
# library(kableExtra)
# library(treemapify)
library(glue)

options(knitr.kable.NA = '--')
knitr::opts_chunk$set(echo=FALSE, message=FALSE)

theme_set(
  theme_bw()
)
```

```{r}
get_pubmed_link <- function(pmid) {
  return(glue::glue("[{pmid}](https://pubmed.ncbi.nlm.nih.gov/{pmid})"))
}
```

```{r}
icite_report = fromJSON(file.path(params$records_file))
search_id = readLines(params$search_id_file)
```

# Introduction

This report shows citation-based and other bibliometric results from [iCite](https://icite.od.nih.gov/) for PRIMED publications.

# All citations

```{r}
icite_analysis_url = glue("https://icite.od.nih.gov/results?search_id={search_id}")
```

View the [full iCite report for all PRIMED publications]( `r icite_analysis_url`).
The full iCite report contains information about citations to PRIMED publications by any other publication.


# Citations within PRIMED only

The PRIMED CC has also generated metrics for citations by PRIMED publications to other PRIMED publications.

```{r}
icite <- icite_report

# Rename icite data frame columns.
names(icite) = names(icite) %>% tolower() %>% str_replace_all(" ", "_")

# Tidy data such that there is one record per "cited_by".
icite <- icite %>%
  unnest_longer(cited_by)
```

```{r}
# Consistency check.
chk <- icite %>%
  group_by(pmid, citation_count) %>%
  summarise(n_cited_by = n()) %>%
  filter(citation_count != n_cited_by)
stopifnot(nrow(chk) == 0)
```

```{r}
# Subset to those that are PRIMED citations.
icite <- icite %>%
  filter(cited_by %in% pmid)
```

## Total citations by publication year

```{r}
# Calculate counts.
#stop()
x = icite %>%
  group_by(year) %>%
  count()
alt_text = paste(
  "Bar chart depicting counts of citations within PRIMED by publication year as follows.",
  x %>% mutate(string=glue("{year}: {n}")) %>% pull(string) %>% paste(collapse = ", ")
)
```

The following figure shows the publication year of the *cited* publication, for PRIMED publications that were cited by another PRIMED publication.

```{r, fig.alt=alt_text}
ggplot(x, aes(x=year, y=n)) +
  geom_bar(stat="identity") +
  geom_text(aes(label = n), vjust = -0.5, size = 3) +
  xlab("Publication year") +
  ylab("Count")
```

## Total citations by year cited

The following figure shows the publication year of the *citing* publication, for PRIMED publications that cited another PRIMED publication.


```{r}
# Calculate counts.
x <- icite %>%
  # First get the distinct set publications citing other publications
  select(cited_by) %>%
  # then get the year those publications were published in
  left_join(icite_report, by=c("cited_by"="pmid")) %>%
  group_by(year) %>%
  count()

alt_text = paste(
  "Bar chart depicting counts of citing publications by publication year as follows.",
  x %>% mutate(string=glue("{year}: {n}")) %>% pull(string) %>% paste(collapse = ", ")
)
```

```{r, fig.alt=alt_text}
ggplot(x, aes(x=year, y=n)) +
  geom_bar(stat="identity") +
  geom_text(aes(label = n), vjust = -0.5, size = 3) +
  xlab("Citation year") +
  ylab("Count")
```

## Most cited publications

The following table shows the 10 PRIMED publications that were most cited by other PRIMED publications.
The "number of citations" columns shows the number of other PRIMED publications citing this publication.

```{r}
most_cited = icite %>%
  group_by(pmid) %>%
  summarise(n_primed_citations=n()) %>%
  arrange(desc(n_primed_citations)) %>%
  head(10)

most_cited %>%
  left_join(icite_report, by="pmid") %>%
  select(pmid, n_primed_citations,  year, title,  authors) %>%
  mutate(pmid=get_pubmed_link(pmid))%>%
  rename(
    `number of citations` = n_primed_citations
  ) %>%
  kable()
```
